Preparing for Potential Closure of European Airspaces due to Re-entering Space Objects

### 2025-06-08 paper.qmd
address reviewer 4 observation on methodology: explained the limitations of ADS-B data usage due to lack of coverage in certain areas and the need to filter out stationary aircraft at airports

The recent decades observed an increase in the number of space activities.
This resulted in higher numbers of re-entries at the end of the respective 
mission lifespan or orbital decay. The risk of collision between uncontrolled
re-entering space debris and aircraft is increasing due to the continual 
growth of flights and of orbiting bodies and their relevant re-entries. 
<!-- why important --> National and international
authorities need to balance safety and economic impact of adopted
measures. This is particularly challenging should a re-entry event be announced on short notice.
Thus, there is a need for effective tools to reduce the uncertainty and increase the safety awareness of uncontrolled re-entries. 
<!-- approach / experiments --> To determine the risk exposure, this paper follows a data-driven approach.
We employ hourly and daily spatial flight densities calculated from trajectories
as recorded by EUROCONTROL's Network Manager and binned with the Uber H3
geospatial indexing system.
Impact risk probabilities are then calculated based on previous theoretical
work. 
<!-- results / conclusions --> This work enhances the previou model by using
actual flown flight trajectories as a basis for time-varying flight densities
rather than relying on proxies such as the population density or ADS-B/Mode S
message counts. These densities combined with a buffered area along the eventual path of re-entry, can provide mid-air impact probabilities to authorities
as an aid for deciding on safety measures such as airspace closures and/or
re-routings.
The findings and the approach are demonstrated on the basis of an expected re-entry from an earlier space mission.\
The approach in this paper enables a computing time efficient mechanism to support the re-entry related decision-making.

@@

We follow the same mathematical approach of [@patera2008;
@wrightAirspaceClosuresDue2025]. However, this paper augments the
approximation of risk exposure. Instead of approximating flight density
with population density [@patera2008] or aircraft transponder data
[@wrightAirspaceClosuresDue2025], we use actual flown flight
trajectories. This provides a more realistic figure of the number of affected air traffic.
Thus, flight density is directly related to the number of flight operations per volume of airspace.

>>>>

The recent decades observed an increase in the number of space activities.
This resulted in higher numbers of re-entries at the end of the respective 
mission lifespan or orbital decay. The risk of collision between uncontrolled
re-entering space debris and aircraft is increasing due to the continual 
growth of flights and of orbiting bodies and their relevant re-entries. 
<!-- why important --> National and international
authorities need to balance safety and economic impact of adopted
measures. This is particularly challenging should a re-entry event be announced on short notice.
Thus, there is a need for effective tools to reduce the uncertainty and increase the safety awareness of uncontrolled re-entries. 
<!-- approach / experiments --> To determine the risk exposure, this paper follows a data-driven approach.
We employ hourly and daily spatial flight densities calculated from trajectories
as recorded by EUROCONTROL's Network Manager and binned with the Uber H3
geospatial indexing system.
Impact risk probabilities are then calculated based on previous theoretical
work. 
<!-- results / conclusions --> This work enhances the previous model by using
actual flown flight trajectories as a basis for time-varying flight densities
rather than relying on proxies such as the population density or ADS-B/Mode S
message counts. These densities combined with a buffered area along the eventual path of re-entry, can provide mid-air impact probabilities to authorities
as an aid for deciding on safety measures such as airspace closures and/or
re-routings.
The findings and the approach are demonstrated on the basis of an expected re-entry from an earlier space mission.\
The approach in this paper enables a computing time efficient mechanism to support the re-entry related decision-making.

@@

We follow the same mathematical approach of [@patera2008;
@wrightAirspaceClosuresDue2025]. However, this paper augments the
approximation of risk exposure. Instead of approximating flight density
with population density [@patera2008] or aircraft transponder data[^adsb]
[@wrightAirspaceClosuresDue2025], we use actual flown flight
trajectories.
This provides a more realistic figure of the number of affected air traffic.
Thus, flight density is directly related to the number of flight operations per 
volume of airspace.

[^adsb]: Crowdsourced ADS-B/Mode S data suffer from lack of coverage away from landmasses and from the need to filter away position reports for stationary aircraft at airports which unnecessarily inflate the density of flying traffic.
